A Theory for Distributed Database Transactions

### 2019-06-14 2018_Transactions/articleTransactions.tex
WIP

\section{Transactions}
\label{sct:Transactions}
	With the theory in place, let us proceed to define transactions.
	Let us first define the notion of action ($\id{act}$)
	and then define the classical properties of transactions: atomic, consistent, isolated, and durable.

	An action is denoted as $\id{act}_{e_1,e_2,\Delta}$, which is a mapping from one DIS to another
\begin{equation}
\label{def:action}
\begin{array}{rl}
	\multicolumn{2}{l}{e_2\not\in\id{events}(\id{prec})\ \wedge\ (e_1,e_2)\in\id{prec}'^*\ \Rightarrow}\\
	&\id{act}_{e_1,e_2,\Delta}(\id{prec},\delta_\id{prec})\ =\ (\id{prec}',\delta'_{\id{prec}'})\\
	&\text{where}\\
	&\ \ \begin{array}[t]{rcll}
		\delta'_{\id{prec}'}(e,e')&=&\Delta&\text{if }(e,e')=(e_1,e_2)\\
		\delta'_{\id{prec}'}(e,e')&=&\delta_\id{prec}(e,e')&\text{otherwise}\\
		\end{array}
\end{array}
\end{equation}
	The state function can be computed recursively as follows:
\begin{equation}
\begin{array}{rcll}
	\id{state}_\id{prec}(e)&=&\id{state}_\id{prec}(e_1)\oplus\delta(e_1,e)&\text{if }(e_1,e)\in\id{prec}\\
	                       &=&\emptyset&\text{otherwise}
\end{array}
\end{equation}

	Now let us define the classical characteristic properties of transactions.
	We start with the easiest one: durability.
	This property expresses that an action must preserve history.
\begin{definition}[Durability]
\item An action $\id{act}_{e_1,e_2,\Delta}$ is durable iff:
\label{def:Durability}
\[\begin{array}{rl}
	&\id{act}_{e_1,e_2,\Delta}(\id{prec},\delta_\id{prec})\ =\ (\id{prec}',\delta'_{\id{prec}'})\\
	\Rightarrow\\
	&\id{prec}^*\subseteq\id{prec}'^*\ \wedge\ \delta_\id{prec}\subseteq\delta'_{\id{prec}'}
\end{array}\]
\end{definition}
	Note that durability is already satisfied in the definition of action (equation~\ref{def:action}). (proof needed)
	Such is not the case with isolation.
	An action $\id{act}_\id{prec}(e_1,e_2)$ is isolated, i.e. performed in isolation, if it does not affect any other action in the system.
	This means that the only event in trace $(e_1,e_2)_\id{prec}$ that precedes events outside that trace is $e_1$.
	Likewise, the only event in trace $(e_1,e_2)_\id{prec}$ that is preceded by events outside that trace is $e_2$.
	To define this, we need an auxiliary function
	$\id{atomize}:\mathbb E\times\mathbb E\times\mathbb H\rightarrow\mathbb H$,
	which replaces a trace $(e_1,e_2)_\id{prec}$ by the pair $(e_1,e_2)$
	in a history $\id{prec}$.
\begin{eqnarray}
	\begin{array}[t]{r@{}l@{}l}
	\multicolumn{3}{l}{\id{atomize}(e_1,e_2,\id{prec})}\\
	   \ =\ \id{prec}-&(&\{ (e_1,e)\in\id{prec}| e_1\ \id{prec}^*\ e\ \wedge\ e\ \id{prec}^*\ e_2\}\ \cup\\
			  & &\{ (e,e_2)\in\id{prec}| e_1\ \id{prec}^*\ e\ \wedge\ e\ \id{prec}^*\ e_2\}\\
			  &)\\
		\cup&&\{ (e_1,e_2) \}\\
	\end{array}
\end{eqnarray}
	We can now define isolation as the property that the traces of the ``atomized'' history
	are the same as the atomization of all traces in that history.
\begin{definition}[Isolation]
\item An action $\id{act}_{e_1,e_2,\Delta}$ works isolated in history $\id{prec}$ iff:
\label{def:Isolation}
\begin{eqnarray}
	(\id{atomize}(e_1,e_2,\id{prec}))^*&=&\id{atomize}(e_1,e_2,\id{prec}^*)
\end{eqnarray}
\end{definition}

	Atomicity is related to isolation because it expresses that a sequence of multiple actions can be treated as one.

\begin{definition}[Atomic]
\item An action $\id{act}_{e_1,e_2,\Delta}$ is atomic in history $\id{prec}$ iff:
\[\begin{array}{rl}
	&\id{act}_{e_1,e_2,\Delta}(\id{prec},\delta_\id{prec})\ =\ (\id{prec}',\delta'_{\id{prec}'})\\
	\Rightarrow\\
	&\id{atomize}(e_1,e_2,\id{prec})\ \leq\ \id{prec}'
\end{array}\]
\label{def:Atomicity}
\end{definition}

	Consistency of an action means that it preserves truth.
\begin{definition}[Consistency]
\item An action $\id{act}:\mathbb H\rightarrow\mathbb H$ is consistent iff it preserves the integrity constraints of all contexts in the system.
\label{def:Consistency}
\begin{eqnarray}
	(h',e',r')=\id{act}(h,e,r)\ \wedge\ r(\id{stat}(h,e))&\Rightarrow&r'(\id{stat}(h',e'))
\end{eqnarray}
\end{definition}

>>>>

\section{Transactions}
\label{sct:Transactions}
	With the theory in place, let us proceed to define transactions.
	Let us first define the notion of transaction
	and then define the classical properties of transactions: atomic, consistent, isolated, and durable.

	A transaction is denoted as $\id{act}_{e_1,e_2,\Delta}$, which is a mapping from one DIS to another
\begin{equation}
\label{def:transaction}
\begin{array}{rl}
	\multicolumn{2}{l}{e_2\not\in\id{events}(\id{prec})\ \wedge\ (e_1,e_2)\in\id{prec}'^*\ \Rightarrow}\\
	&\id{act}_{e_1,e_2,\Delta}(\id{prec},\delta_\id{prec})\ =\ (\id{prec}',\delta'_{\id{prec}'})\\
	&\text{where}\\
	&\ \ \begin{array}[t]{rcll}
		\delta'_{\id{prec}'}(e,e')&=&\Delta&\text{if }(e,e')=(e_1,e_2)\\
		\delta'_{\id{prec}'}(e,e')&=&\delta_\id{prec}(e,e')&\text{otherwise}\\
		\end{array}
\end{array}
\end{equation}
	The state function can be computed recursively as follows:
\begin{equation}
\begin{array}{rcll}
	\id{state}_\id{prec}(e)&=&\id{state}_\id{prec}(e_1)\oplus\delta(e_1,e)&\text{if }(e_1,e)\in\id{prec}\\
	                       &=&\emptyset&\text{otherwise}
\end{array}
\end{equation}

	Now let us define the classical characteristic properties of transactions.
	Let us start with atomic.
	Atomicity expresses that a trace of multiple events can be treated as a single pair $\in\id{prec}$.
	To define this, we need an auxiliary function
	$\id{atomize}:\mathbb E\times\mathbb E\times\mathbb H\rightarrow\mathbb H$,
	which replaces a trace $(e_1,e_2)_\id{prec}$ by the pair $(e_1,e_2)$
	in a history $\id{prec}$.
\begin{eqnarray}
	\begin{array}[t]{r@{}l@{}l}
	\multicolumn{3}{l}{\id{atomize}(e_1,e_2,\id{prec})}\\
	   \ =\ \id{prec}-&(&\{ (e',e)\in\id{prec}| e_1\ \id{prec}^*\ e\ \wedge\ e\ \id{prec}^*\ e_2\}\ \cup\\
			  & &\{ (e,e')\in\id{prec}| e_1\ \id{prec}^*\ e\ \wedge\ e\ \id{prec}^*\ e_2\}\\
			  &)\\
		\cup&&\{ (e_1,e_2) \}\\
	\end{array}
\end{eqnarray}

\begin{definition}[Atomic]
\item A transaction $\id{act}_{e_1,e_2,\Delta}$ is atomic in history $\id{prec}$ iff:
\[\begin{array}{rl}
	&\id{act}_{e_1,e_2,\Delta}(\id{prec},\delta_\id{prec})\ =\ (\id{prec}',\delta'_{\id{prec}'})\\
	\Rightarrow\\
	&\id{atomize}(e_1,e_2,\id{prec})\ \leq\ \id{prec}'
\end{array}\]
\label{def:Atomicity}
\end{definition}

	Now let us turn to consistency.
	This property means that a transaction preserves truth of the state.
	We need a predicate $\mathcal P\{\mathbb S\}$ to express that a state is consistent.
\begin{definition}[Consistency]
\item A transaction $\id{act}_{e_1,e_2,\Delta}$ is consistent w.r.t. $\mathcal P$ iff:
\[\mathcal P\{\id{state}_\id{prec}(e_1)\}\ \Rightarrow\ \mathcal P\{\id{state}_\id{prec}(e_2)\}\]
\label{def:Consistency}
\end{definition}

	The third property is isolation.
	A transaction $\id{act}_\id{prec}(e_1,e_2)$ is isolated, i.e. performed in isolation, if it does not affect any other transaction in the system.
	This means that the only event in trace $(e_1,e_2)_\id{prec}$ that precedes events outside that trace is $e_1$.
	Likewise, the only event in trace $(e_1,e_2)_\id{prec}$ that is preceded by events outside that trace is $e_2$.
	We can now define isolation as the property that the traces of the ``atomized'' history
	are the same as the atomization of all traces in that history.
\begin{definition}[Isolation]
\item A transaction $\id{act}_{e_1,e_2,\Delta}$ works isolated in history $\id{prec}$ iff:
\label{def:Isolation}
\begin{eqnarray}
	(\id{atomize}(e_1,e_2,\id{prec}))^*&=&\id{atomize}(e_1,e_2,\id{prec}^*)
\end{eqnarray}
\end{definition}

	Then we have durability.
	This property expresses that a transaction must preserve history.
\begin{definition}[Durability]
\item A transaction $\id{act}_{e_1,e_2,\Delta}$ is durable iff:
\label{def:Durability}
\[\begin{array}{rl}
	&\id{act}_{e_1,e_2,\Delta}(\id{prec},\delta_\id{prec})\ =\ (\id{prec}',\delta'_{\id{prec}'})\\
	\Rightarrow\\
	&\id{prec}^*\subseteq\id{prec}'^*\ \wedge\ \delta_\id{prec}\subseteq\delta'_{\id{prec}'}
\end{array}\]
\end{definition}
	Note that durability is already satisfied in the definition of transaction (equation~\ref{def:transaction}). (proof needed)
